---
title: "County Partisan Leanings"
author: "Aman Choudhri"
date: "2024-10-01"
output: pdf_document
---

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
```

```{r}
setwd("C:/Users/jared/Desktop/Jared/social-ties")
```

```{r}
precinct_data <- read.csv('dat/mit-edsl-prez-by-precinct-2020/PRESIDENT_precinct_general.csv')
```

Get Democratic and Republican voteshares.
```{r}
# map Minnesota's Democratic Farmer Labor party to listing as Democrat
precinct_data <- precinct_data %>% mutate(party_simplified=ifelse(
  state == "MINNESOTA" & party_detailed=="DEMOCRATIC FARMER LABOR",
  "DEMOCRAT",
  party_simplified
  ))
dr_totals_only = precinct_data[precinct_data$party_simplified %in% c("DEMOCRAT", "REPUBLICAN"),]
dr_totals_only
```

Aggregate by county.

```{r}
totals_by_county <- dr_totals_only %>% group_by(state, county_name, party_simplified) %>% summarize(n=sum(votes))
totals_by_county
```
Now get the vote share proportions.
```{r}
by_county <- totals_by_county %>% mutate(total_votes=sum(n)) %>% summarize(
  democrat_votes=n[party_simplified=="DEMOCRAT"],
  total_votes=first(total_votes),
  proportion_democrat=democrat_votes/total_votes,
  .groups="drop"
)
by_county
```
Process any naming issues with `by_county`.
```{r}
by_county[(by_county$county_name == "ST LOUIS COUNTY") & (by_county$state == "MISSOURI"), "county_name"] <- "ST LOUIS"
```
Drop the entry for `KANSAS CITY` since those votes are already counted in `JACKSON` county.
Also, throw out overseas voters in Maine, since we don't know what county they are from.
And delete a blank row for Rhode Island with no county information.
```{r}
is_kcmo <- (by_county$state == "MISSOURI") & (by_county$county_name == "KANSAS CITY")
is_uocava <- (by_county$state == "MAINE") & (by_county$county_name == "STATEWIDE UOCAVA")
is_ri_state_only<- (by_county$state == "RHODE ISLAND") & (by_county$county_name == "")
by_county <- by_county %>% filter(!is_kcmo & !is_uocava)
```

```{r}
ggplot(by_county, aes(x=proportion_democrat)) + geom_histogram() + ggtitle('County Democratic Voteshares in 2020 Election')
```

Calculate tercile sizes, if we say a `swing` county is between 0.45 and 0.55.
```{r}
lower_bound = 0.45
upper_bound = 0.55
by_county$tercile <- ifelse(
  by_county$proportion_democrat < lower_bound, 1,
  ifelse(
    by_county$proportion_democrat < upper_bound, 0, -1
  )
)

label <- c(
  paste0("prop_dem < ", lower_bound),
  paste0(lower_bound, " < prop_dem < ", upper_bound),
  paste0("prop_dem > ", upper_bound)
)
by_county$tercile_label <- factor(label[by_county$tercile + 2], levels=label)
by_county
```

Get the populations of each county.
```{r}
# Data: US census
# https://www.census.gov/data/tables/time-series/demo/popest/2020s-counties-total.html

populations_raw <- read.csv('dat/DECENNIALPL2020.P1-Data.csv', header=T, skip=1)

# basic QoL variable renaming and selection
populations <- populations_raw %>%
  rename(geographic_area_name = Geographic.Area.Name, population=X...Total. ) %>%
  select(geographic_area_name, population)

populations
```
```{r}
geo_info <- as_tibble(populations$geographic_area_name %>% str_split_fixed(", ", n=2))
colnames(geo_info) <- c("county_raw", "state_raw")
geo_info
```
```{r}
populations <- populations %>% bind_cols(geo_info)
populations
```

```{r}
# separate county (or census area) and state
# match formatting of state and county names from the MIT dataset
populations$state <- populations$state_raw %>% str_to_upper()
populations$county_name <- (populations$county_raw %>%
  str_to_upper() %>%
  str_remove(" COUNTY") %>%
  str_remove(" PARISH") %>%
  str_replace("^ST\\. ", "ST ") %>%
  str_replace("^STE\\. ", "STE ")
)
# handle naming quirks
populations[(populations$state == "NEW MEXICO") & (populations$county_name == "DOÑA ANA"), "county_name"] <- "DONA ANA"
populations[(populations$state == "MISSOURI") & (populations$county_name == "DEKALB"), "county_name"] <- "DE KALB"
populations[(populations$state == "LOUISIANA") & (populations$county_name == "LASALLE"), "county_name"] <- "LA SALLE"
# populations[(populations$county_name == "ST LOUIS CITY") & (populations$state == "MISSOURI"), "county_name"] <- "ST LOUIS"
# select only the population estimate from 2020
pops_to_merge <- populations %>% select(c(state, county_name, population))
pops_to_merge
```

Merge them back into our county data.
```{r}
by_county_merged <- merge(
  by_county,
  pops_to_merge,
  by.x = c('state', 'county_name'),
  by.y = c('state', 'county_name'),
  all.x=T
)
by_county_merged
```
```{r}
by_county_merged[is.na(by_county_merged$population),]
```
Lastly, aggregate all populations across subunits in Alaska and Rhode Island, since
we don't have subunit-level voting information in those areas.
```{r}
state_level_pop <- populations %>%
  group_by(state) %>%
  summarize(population = sum(population)) %>%
  filter(state %in% c("ALASKA", "RHODE ISLAND"))

by_county_merged <- by_county_merged %>%
  left_join(state_level_pop, by = "state") %>%
  mutate(population = coalesce(population.x, population.y)) %>%
  select(-population.x, -population.y)

```

Check to make sure we have all of our population information.
```{r}
sum(is.na(by_county_merged$population))
```

```{r}
by_county_merged
```
```{r}
state_level_pop
```

```{r}
by_county_merged %>% group_by(tercile_label) %>% summarize(
  n_counties = n(),
  n_votes = sum(total_votes),
  n_people = sum(population)
)
```


```{r}
out <- by_county_merged %>% select(state, county_name, population, total_votes, proportion_democrat, tercile)
write.csv(out, 'dat/county-partisan-leanings.csv')
```