---
title: "neighbor-polling-2"
author: "Jared Winslow"
date: "2024-12-10"
output: html_document
---

# Neighbor Polling Exploration


```{r, echo=T, message=FALSE}
library(dplyr)
library(here)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(here::here())
```

```{r}
# notebook-level config parameters
# whether averages should
# be calculated using survey weights or not
use_weights <- TRUE
```

```{r}
# load in the wave 10 dataframe
df <- readRDS('dat/wave6/processed_wave_6.rds')

if (!use_weights) {
  df$weight <- 1
}
```

## "Friend Group" Polling

### Exploratory Analysis
```{r}
politics_counts <- table(df$friend_group_discuss_politics)
table(df$friend_group_discuss_politics, df$friend_group_pid5)
```

### Missing Data Plan

Throw out 1. people who don't respond to `friend_group_presvote24`, 2.
those who report not planning on voting.

### Experiment 1: Estimating Friend Group Polling Adjustment

To estimate the "popular vote" going to Trump under friend group polling, we'll
take advantage of the `friend_group_presvote24` variable, which asked:

> Who do you think most of your friends will vote for in the presidential election
> in November?

With answer choices of:

 * Mostly Kamala Harris
 * About evenly split
 * Mostly Donald Trump
 * Mostly would not vote
 * Mostly another candidate
 * Not sure

For this analysis, we'll discard everyone who thinks their friends
won't vote, will vote third party, or aren't sure.

```{r}
knitr::kable(
  table(df$friend_group_presvote24),
  col.names=c("Friend Group Vote Estimate", "Count")
  )
```


```{r}
include <- !is.na(df$friend_group_presvote24) & (
  df$friend_group_presvote24 == "Mostly Democrat" |
  df$friend_group_presvote24 == "About evenly split" |
  df$friend_group_presvote24 == "Mostly Republican"
)

print(paste('This excludes', sum(1 - include, na.rm=T), 'people.'))
```

Let's code friend group presidential vote by considering a "Mostly" response
to mean 75%. We can ablate on this later.

```{r}
mostly_meaning <- 0.75

codebook <- c(
  "Mostly Democrat" = 1 - mostly_meaning,
  "About evenly split" = 0.5,
  "Mostly Republican" = mostly_meaning
)

df$friend_group_presvote24_coded <- codebook[df$friend_group_presvote24]

knitr::kable(
  table(df$friend_group_presvote24_coded),
  col.names=c("Estimated Friends Voteshare for Trump", "Count")
  )
```


Now just compute the difference between the average friend group polling result
and the average standard polling result. To compute the standard polling result,
we'll similarly include only individuals report planning to vote either for
Harris or Trump.

```{r}
standard_estimate <- df %>%
  
  mutate(coded_vote = (presvote24 == "Republican")) %>%
  summarise(est=mean(coded_vote * weight)) %>%
  pull(est)

friend_group_estimate <- mean(df$friend_group_presvote24_coded * df$weight)

display_digits <- 4
print(paste(
  "Result with standard polling:",
  round(standard_estimate, display_digits)
  ))
print(paste(
  "Result with friend group polling:",
  round(friend_group_estimate, display_digits)
  ))
```

Now we calculate the standard errors of the estimates.

```{r}
library(Hmisc)

df <- df %>%
  mutate(coded_vote = (presvote24 == "Republican"))
  
individual_se <- sqrt(wtd.var(df$coded_vote, df$weight) / sum(df$weight))

friend_se <- sqrt(
  wtd.var(df$friend_group_presvote24_coded, df$weight)
  / sum(df$weight)
  )

print(paste(
  "SE of standard polling:",
  round(individual_se, display_digits)
  ))
print(paste(
  "SE of friend group polling:",
  round(friend_se, display_digits)
  ))
```

### Experiment 1.5: Subset of Respondents
Since we calculate the `individual` estimate and the `friend_group` estimates
from different subsets, we don't have a way to estimate the standard error of
the difference between the two estimates.

In this section, we'll recalculate the estimates from the same subset of
respondents: major-party voters who report their friends as mostly major-party
voters.

Now let's renormalize the survey weights.
```{r}
df$weight <- df$weight / mean(df$weight)
```

```{r}
subset_individual_estimate <- df %>%
  mutate(coded_vote = (presvote24 == "Republican")) %>%
  summarise(est=mean(coded_vote * weight)) %>%
  pull(est)

subset_friend_estimate <- mean(
  df$friend_group_presvote24_coded *
  df$weight
  )

cat(
  "Subset estimates ----",
  "Results without subsetting in parenthesis.",
  paste0(
    "\nStandard polling: ",
    round(subset_individual_estimate, display_digits)
    #,
    # " (",
    # round(standard_estimate, display_digits),
    # ")"
  ),
  paste0(
    "\nFriend group polling: ",
    round(subset_friend_estimate, display_digits)
    #,
    #" (", round(friend_group_estimate, display_digits), ")"
  ),
  paste(
    "\nSubset difference in standard polling:",
    round(subset_individual_estimate - standard_estimate, display_digits)
  ),
  paste(
    "\nSubset difference in friend group polling:",
    round(subset_friend_estimate - friend_group_estimate, display_digits)
  ),
  sep="\n"
  )
```

Now we can calculate the standard error of the difference between friend group
polling and individual polling.

```{r}
diffs <- df %>%
  mutate(d = friend_group_presvote24_coded - coded_vote) %>%
  pull(d)

se <- sqrt(
  wtd.var(diffs, df$weight)
  / sum(df$weight)
  )

print(paste("SE:", round(se, display_digits)))
```

This gives the following 95% confidence interval on the difference.

```{r}
diff <- subset_friend_estimate - subset_individual_estimate
cat(
  round(diff, display_digits), "+/- 2se:",
  paste0(
    "[",
    round(diff - (2 * se), display_digits),
    ", ",
    round(diff + (2 * se), display_digits),
    "]"
  ),
  sep=" "
)
```
