---
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

# Neighbor Polling Exploration Wave 10

```{r, echo=T, message=FALSE}
library(dplyr)
library(here)
library(Hmisc)
library(tidyr)
library(arm)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(here::here())
```

```{r}
# notebook-level config parameters
# whether averages should
# be calculated using survey weights or not
use_weights <- TRUE
```

```{r}
print_shape <- function(df) {
  cat("Rows:", nrow(df), "| Columns:", ncol(df), "\n")
  df}
```

## "Friend Group" Polling

### Missing Data Plan

Ignore responses from people who didn't respond to `presvote24h` or
`friend_group_presvote24`. Also, we'll throw out responses from
non-major party voters. Specifically, those who report:

-   Planning to vote third party,
-   Planning not to vote, or
-   Being unsure about their vote.

```{r}
# load in the wave 10 dataframe
df <- readRDS('dat/wave10/processed_wave_10.rds')

if (!use_weights) {
  df$weight <- 1
}

initial_count <- nrow(df)
df <- df %>% 
  print_shape() %>%
  # Eliminate responses with missing values
  drop_na(presvote24h, friend_group_presvote24h) %>% 
  print_shape() %>%
  # Restrict to major-party voters
  filter(presvote24h %in% c("Democrat", "Republican")) %>%
  print_shape()
n_excluded <- initial_count - nrow(df)
print(paste('This excludes', n_excluded, 'people.'))
```

Renormalize the survey weights.

```{r}
df$weight <- df$weight / mean(df$weight)
```

### Exploratory Analysis

```{r}
politics_counts <- table(df$friend_group_discuss_politics)
table(df$friend_group_discuss_politics, df$friend_group_pid5)
```

### Experiment 1: Estimating Friend Group Polling Adjustment

To estimate the "popular vote" going to Trump under friend group
polling, we'll take advantage of the `friend_group_presvote24` variable,
which asked:

> Who do you think most of your friends will vote for in the
> presidential election in November?

With answer choices:

-   Mostly Kamala Harris
-   About evenly split
-   Mostly Donald Trump
-   Mostly would not vote
-   Mostly another candidate
-   Not sure

We collapsed the latter three answer choices into "About evenly split,"
since they're not clearly partisan one way or another. See the
`recode_presvote` method in `process-data.R` for more details.

```{r}
knitr::kable(
  table(df$friend_group_presvote24h),
  col.names=c("Friend Group Vote Estimate", "Count")
  )
```

Let's code friend group presidential vote by considering a "Mostly"
response to mean 75%. We perform a sensitivity analysis on this later;
the exact choice turns out not to matter much.

```{r}
mostly_meaning <- 0.75

codebook <- c(
  "Democrat" = 1 - mostly_meaning,
  "About evenly split" = 0.5,
  "Republican" = mostly_meaning
)

df$friend_group_presvote24h_coded <- codebook[df$friend_group_presvote24h]

knitr::kable(
  table(df$friend_group_presvote24h_coded),
  col.names=c("Estimated Friends Voteshare for Trump", "Count")
  )
```

Now just compute the difference between the average friend group polling
result and the average standard polling result. To compute the standard
polling result, we'll similarly include only individuals report planning
to vote either for Harris or Trump.

```{r}
df <- df %>%
  mutate(coded_vote = (presvote24h == "Republican"))

standard_estimate <- df %>%
  dplyr::select(coded_vote, weight) %>%
  summarise(est=mean(coded_vote * weight)) %>%
  pull(est)

friend_group_estimate <- mean(df$friend_group_presvote24h_coded * df$weight)

display_digits <- 4
print(paste(
  "Result with standard polling:",
  round(standard_estimate, display_digits)
  ))
print(paste(
  "Result with friend group polling:",
  round(friend_group_estimate, display_digits)
  ))
```

Now we calculate the standard errors of the estimates.

```{r}
individual_se <- sqrt(wtd.var(df$coded_vote, df$weight) / sum(df$weight))

friend_se <- sqrt(
  wtd.var(df$friend_group_presvote24h_coded, df$weight)
  / sum(df$weight))

print(paste(
  "SE of standard polling:",
  round(individual_se, display_digits)))

print(paste(
  "SE of friend group polling:",
  round(friend_se, display_digits)))
```

Last, calculate the standard error of the difference between friend
group polling and individual polling.

```{r}
diffs <- df %>%
  mutate(d = friend_group_presvote24h_coded - coded_vote) %>%
  pull(d)

se <- sqrt(
  wtd.var(diffs, df$weight)
  / sum(df$weight)
  )

print(paste("SE:", round(se, display_digits)))
```

This gives the following 95% confidence interval on the difference.

```{r}
diff <- friend_group_estimate - standard_estimate
cat(
  round(diff, display_digits), "+/- 2se:",
  paste0(
    "[",
    round(diff - (2 * se), display_digits),
    ", ",
    round(diff + (2 * se), display_digits),
    "]"
  ),
  sep=" "
)
```

### Experiment 2: Deviations from PID

Next, we'll compare these polling estimates to the average individual
reported party ID and average estimated friend group party ID.

```{r}
# for now, throw out Independent/Not sure respondents
individual_pid <- df %>%
  filter(collapsed_pid != "Independent") %>%
  dplyr::summarize(est=mean((collapsed_pid == "Republican") * weight)) %>%
  pull(est)
```

To compute the friend group PID estimate, we'll throw out responses from
people who reported that they aren't sure about the party composition of
their friends.

```{r}
codebook <- c(
  "All Democrats"= 0,
  "Mostly Democrats" = 1 - mostly_meaning,
  "About evenly split" = 0.5,
  "Mostly Republicans" = mostly_meaning,
  "All Republicans" = 1
  )

friend_group_pid <- df %>%
  filter(friend_group_pid5 != "Not sure") %>%
  mutate(friend_group_pid5_coded=codebook[friend_group_pid5]) %>%
  dplyr::summarize(est=mean(friend_group_pid5_coded * weight)) %>%
  pull(est)
```

```{r}
display_digits <- 4
print(paste(
  "Proportion Republican with standard polling:",
  round(individual_pid, display_digits)
  ))
print(paste(
  "Proportion Republican with friend group polling:",
  round(friend_group_pid, display_digits)
  ))
```

Now we can compute the differences.

```{r}
standard_estimate - individual_pid
```

```{r}
friend_group_estimate - friend_group_pid
```

### Experiment 3: Sensitivity Analysis

Let's check how sensitive this 'neighbor bias' is to the choice of
coding we use for "Mostly ...".

```{r}
# get the average survey weight per response category,
# weighted by the size of the response category
vote_weights <- df %>%
  group_by(friend_group_presvote24h) %>%
  dplyr::summarize(weight=mean(weight), n=n()) %>%
  mutate(p = n / sum(n), weight=p*weight) %>%
  dplyr::select(friend_group_presvote24h, weight) %>%
  tibble::deframe()

vote_weights
```

```{r}
friend_group_poll_est <- function (p) {
  return (
    (1 - p) * vote_weights["Democrat"] +
    0.5 * vote_weights["About evenly split"] +
    p * vote_weights["Republican"]
  )
}

mostly_code_options <- seq(0.5, 1, length.out=100)
plot(
  mostly_code_options,
  friend_group_poll_est(mostly_code_options),
  main="Trump Voteshare Poll Estimate by Proportion Parameter",
  xlab="p",
  ylab="Trump Vote Share",
  )
```

```{r}
pid_weights <- df %>%
  group_by(friend_group_pid5) %>%
  dplyr::summarize(weight=mean(weight), n=n()) %>%
  mutate(p = n / sum(n), weight=p*weight) %>%
  dplyr::select(friend_group_pid5, weight) %>%
  tibble::deframe()

friend_group_pid_est <- function (p) {
  return (
    0 * pid_weights['All Democrats'] +
    (1 - p) * pid_weights["Mostly Democrats"] +
    0.5 * pid_weights["About evenly split"] +
    p * pid_weights["Mostly Republicans"] +
    1 * pid_weights["All Republicans"]
  )
}

plot(
  mostly_code_options,
  friend_group_pid_est(mostly_code_options),
  main="Republican Composition Estimate by Proportion Parameter",
  xlab="p",
  ylab="Republican Composition",
  )
```

Poll adjustment.

```{r}
poll_adjustments <- (
  friend_group_poll_est(mostly_code_options) -
    friend_group_pid_est(mostly_code_options)
  )
plot(
  mostly_code_options,
  poll_adjustments,
  main="Friend Group Poll Adjustment by Proportion Parameter",
  xlab="p",
  ylab="Increase in Trump Voteshare with Friend Poll",
  )
```

Even in the worst case, the friend group poll estimate is correctly
pulled in favor of Trump.

```{r}
poll_adjustments[length(poll_adjustments)]
```

### Experiment 4: Individual-Level Analysis

For each person, we'll compute the following differences:

$$
\text{pid}_\text{self} - \text{pid}_\text{friends}, \quad \text{and} \quad
\text{vote}_\text{self} - \text{vote}_\text{friends}.
$$

Comparing these differences can help us get a more granular sense for
the efficacy of the neighbor polling methodology.

For this analysis, we'll encode our variables numerically on a 0-1 scale to
stay consistent with the previous analyses.

```{r}
to_zero_one <- function(var) {
  # Convert a [-1, 1] variable to [0, 1].
  return(0.5 * (1 + var))
}

encode_to_zero_one <- function(df) {
  df %>% mutate(across(
    c(collapsed_pid_numeric, friend_group_pid_numeric,
      presvote24h_numeric, friend_group_presvote24h_numeric),
    to_zero_one
  ))
}

df <- df %>% encode_to_zero_one()
```

Quick sanity test.
```{r}
par(mfrow=c(2,2))

hist(df$collapsed_pid_numeric)
hist(df$friend_group_pid_numeric)
hist(df$presvote24h_numeric)
hist(df$friend_group_presvote24h_numeric)
```

```{r}
individual_deviations <- df$presvote24h_numeric - df$collapsed_pid_numeric
friend_deviations <- df$friend_group_presvote24h_numeric - df$friend_group_pid_numeric

jitter <- rnorm(length(individual_deviations), 0, 0.003)

plot(individual_deviations + jitter, friend_deviations)
```

```{r}
g <- lm(friend_deviations ~ individual_deviations)
display(g)
```

```{r}
plot(individual_deviations + jitter, friend_deviations)
abline(g, col='blue')
```


## "Best Friend" Polling

### Experiment 1: Estimating Best Friend Polling Adjustment

For another estimate of the "popular vote" going to Trump under neighbor
polling, we'll take advantage of the `best_friend_group_presvote24`
variable from wave 6, which asked:

> Who do you think [your closest friend] will vote for in the
> presidential election in November?

With answer choices:

-   Joe Biden
-   Donald Trump
-   Robert F. Kennedy, Jr.
-   Jill Stein
-   Cornel West
-   Other
-   Would not vote
-   Not sure

For a first analysis, we'll discard everyone who reported their best
friends would vote for anyone other than Biden or Trump.

```{r}
# df <- readRDS('dat/week-6/processed.rds')
# include <- (
#   df$best_friend_presvote24 == "Joe Biden" |
#   df$best_friend_presvote24 == "Donald Trump"
# )
# df <- df[include,]
```

From this, we can simply calculate the difference in means of reported
2024 vote among
