---
author: "Aman Choudhri"
date: "2024-10-22"
params:
  img_dir:
    value: 'figs/regression/'
  base_dir:
    value: '/Users/amanchoudhri/aman/penumbra/notebooks/'
  output_mode:
    value: 'pdf'
output:
  github_document:
    html_preview: false
  pdf_document: 
    fig_height: 5
title: "Regressing Vote on Covariates"
editor_options: 
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dpi=300,fig.width=7)

if (params$output_mode == 'github') {
  base.dir <- '/Users/amanchoudhri/aman/penumbra/'
  base.url <- '/'
  
  knitr::opts_knit$set(base.dir=params$base_dir)
  knitr::opts_chunk$set(fig.path = params$img_dir)
}
```

```{r libraries, echo=T, message=FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
library(patchwork)

library(arm)
library(rstanarm)
```

In this notebook, we regress individuals' votes on covariates like their
individual partisan leanings, the partisan leanings of their state/county, the
copartisanship of their friend groups, etc.

First set up our data.

```{r}
CLEANED_DATA_FILENAME <- 'dat/processed.rds'

df <- readRDS(CLEANED_DATA_FILENAME)
```

Load in county-level voteshares from 2020.
```{r}
county_leanings <- read.csv('dat/county-partisan-leanings.csv')
county_leanings <- county_leanings %>% rename(county_p_dem=proportion_democrat)
knitr::kable(
  county_leanings %>%
    mutate(state=str_to_sentence(state), county_name=str_to_sentence(county_name)) %>%
    slice_head(n=5),
  digits=3
  )
```
Also aggregate up and calculate state leanings.
```{r}
state_leanings <- county_leanings %>%
  group_by(state) %>%
  mutate(frac_state_votes=total_votes/sum(total_votes)) %>%
  summarize(state_p_dem = sum(county_p_dem * frac_state_votes))

knitr::kable(
  state_leanings %>%
    mutate(state=str_to_sentence(state)) %>%
    slice_head(n=10),
  digits=3
  )
```

And merge both into `df`.
```{r}
df <- df %>% mutate(state_upper = toupper(inputstate), county_upper=toupper(county)) %>%
  left_join(
    county_leanings,
    by=c("state_upper"="state", "county_upper"="county_name")
  ) %>%
  left_join(
    state_leanings,
    by=c("state_upper"="state")
  ) %>% mutate(
    state_p_rep = 1 - state_p_dem,
    county_p_rep = 1 - county_p_dem
  )
```

```{r, display-merged-table, eval=T, echo=F}
display_df <- function(data) {
  knitr::kable(
    data %>%
      mutate(state=str_to_sentence(inputstate)) %>%
      mutate(
        party_id=case_when(
          collapsed_pid == "Democrat" ~ "D",
          collapsed_pid == "Republican" ~ "R",
          .default = "I"
        ),
        vote_in_2020=case_when(
          presvote20post == "Joe Biden" ~ "D",
          presvote20post == "Donald Trump" ~ "R",
          .default = "Other/NA"
        )
      ) %>%
      dplyr::select(party_id, vote_in_2020, state, county, state_p_rep, county_p_rep) %>%
      slice_head(n=10),
    digits=3
    )
}
```

```{r}
display_df(df)  # function definition in Appendix
```

For the regressions, we'll convert the `friend_group_pid5` variable into a
numeric score from -2 to 2 (All Democrats, ..., All Republicans). We'll collapse
respondents who marked "Not sure" into the same bucket as those who marked
"About evenly split".
```{r}
df <- df %>% mutate(
  friend_group_pid = ifelse(friend_group_pid5 == "Not sure", 0, as.numeric(friend_group_pid5) - 3),
)
df %>% dplyr::select(friend_group_pid5, friend_group_pid)
```

## Predicting 2020 Vote
For simplicity, restrict to only people who voted in 2020, and only those who voted for
Trump/Biden.

```{r}
print(paste0(
  '# of people who didn\'t vote in 2020: ',
  nrow(df %>% filter(presvote20post == "Did not vote for President"))
))
voted_d_or_r <- 
print(paste0(
  '# of people who voted third-party in 2020: ',
  nrow(df %>% filter(
    presvote20post != "Did not vote for President",
    !(presvote20post %in% c("Donald Trump", "Joe Biden"))
    )
    )
))
```
```{r}
restricted <- df %>%
  filter(presvote20post %in% c("Donald Trump", "Joe Biden")) %>%
  mutate(
    vote = ifelse(presvote20post == "Donald Trump", 1, 0),
    pid = as.numeric(collapsed_pid) - 2,
    state_p_rep = 1 - state_p_dem,
    county_p_rep = 1 - county_p_dem,
  )
knitr::kable(
  restricted %>%
    dplyr::select(
      pid, friend_group_pid, vote, inputstate,
      county, state_p_rep, county_p_rep
      ) %>% 
    rename(state=inputstate) %>%
    slice_head(n=5),
  digits=3
)
```

### Model 1: State Leaning

```{r}
model_1 <- stan_glm(vote ~ state_p_rep, data=restricted, family=binomial, refresh=0)
```

```{r}
summary(model_1)
```
Interpreting these coefficients using the "divide by 4 trick", we learn that a
0.1-unit Republican shift in state vote share will in expectation raise the probability
of an individual voting Republican by 0.38 / 4 = 0.095.

### Model 2: County
```{r}
model_2 <- stan_glm(vote ~ state_p_rep + county_p_rep, data=restricted, family=binomial, refresh=0)
```

```{r}
summary(model_2)
```
Interpreting these coefficients, it seems that state Republican voteshare has a
much more modest relationship with an individual's propensity to vote
Republican, after adjusting for county information. Now, a 10 percentage point
increase in Republican voteshare will in expectation mean an increase
of 1.75 percentage points in the probability of voting Republican.

The much stronger correlation, as expected is with county Republican voteshare.
Repeating the same divide by four trick, a 10pp increase in county Republican
voteshare will in expectation correspond to a 8.25pp increase in probability of
having voted Republican in 2020.

### Model 3: Friend Group Partisanship
```{r}
model_3 <- stan_glm(
  vote ~ state_p_rep + county_p_rep + friend_group_pid,
  data=restricted, family=binomial, refresh=0
  )
```

```{r}
summary(model_3)
```

As expected, the state-level information adds minimal predictive power.
Interestingly, county Republican voteshare still seems to predict vote after
adjusting for friend group partisanship. A 10pp Republican shift in county
voteshare will in expectation result in a 2.5pp increase in the probability of
an individual having voted Republican.

With friend group partisanship, the interpretation of the coefficient is that a
one-unit Republican shift in friend group partisan composition (i.e. from "All
Democrats" to "Mostly Democrats") will *at maximum* correspond to a whopping
52pp increase in probability of voting Republican. This figure is primarily
meaningful around the center, representing shifts in friend group partisanship between
"Mostly Democrat", "About evenly split", and "Mostly Republican".

### Model 4: Individual Party ID
Finally, incorporate individual party ID into the model.
```{r}
model_4 <- stan_glm(
  vote ~ state_p_rep + county_p_rep + friend_group_pid + pid,
  data=restricted, family=binomial, refresh=0
  )
```

```{r}
summary(model_4)
```

A one-unit rightward shift in `pid` (from Democrat to Independent/Other or from
I/O to Republican) will at maximum lead to an expected increase in probability
of voting Republican by 75pp. After adjusting for this, however, friend group
partisanship still seems to be strongly predictive of vote, with a 1-unit
rightward shift in friend group partisan composition leading to a maximum of a
22.5pp increase in probability of voting Republican.

### Model 5: More granular personal PID

Here we check the intuition that `friend_group_pid5` is maybe just a proxy for
strength of personal party identification. If it is, we might expect that adding
a more granular signal of personal party ID would remove some predictive power from
the friend group covariate.

```{r}
restricted <- restricted %>% mutate(
  granular_pid=ifelse(
    pid7 == "Not sure",
    0,
    as.numeric(pid7) - 4
  )
)

knitr::kable(
  restricted %>% dplyr::select(pid7, granular_pid) %>% slice_head(n=10)
  )
```
```{r}
model_5 <- stan_glm(
  vote ~ pid + friend_group_pid + state_p_rep + county_p_rep + granular_pid,
  data=restricted,
  family=binomial,
  refresh=0
)
```
```{r}
summary(model_5)
```
Why would `pid` still have some predictive power under this model? The `pid`
variable is just a collapsed version of `granular_pid`—in fact, `pid` shouldn't
add any information beyond `granular_pid` whatsoever. Conceptually this might
result in collinearity and hence inflated posterior standard error, but it
doesn't really look like we see that.

Let's visualize the posterior distribution on $\beta_{\text{pid}}$.
```{r}
draws_4 <- as.data.frame(model_4$stanfit)
draws_5 <- as.data.frame(model_5$stanfit)

draws <- bind_rows(list('m4'=draws_4, 'm5'=draws_5), .id="id")
```

```{r}
ggplot(draws, aes(pid)) + geom_histogram() + facet_grid(. ~ id)
```

Let's try without `pid` for more interpretability.

### Model 6: Granular PID only

```{r}
model_6 <- stan_glm(
  vote ~ friend_group_pid + state_p_rep + county_p_rep + granular_pid,
  data=restricted,
  family=binomial,
  refresh=0
)
```

```{r}
summary(model_6)
```


### Multilevel Structure
```{r}
multilevel <- stan_glmer(
  vote ~ pid + friend_group_pid + state_p_rep + county_p_rep + (1 | inputstate),
  data=restricted,
  family=binomial
  )
```

```{r}
summary(multilevel)
```
## Appendix
### Utility functions
```{r, display-df, eval=F}
```

### Model diagnostics
